A Triple Origin for Twin Blue Stragglers in Close Binaries

### 2019-03-25 BSS7782_prev.tex
added previous version to the git.

We propose a formation mechanism for twin blue stragglers (BSs) in
compact binaries that involves mass transfer from an evolved outer
tertiary companion on to the inner binary via a circumbinary disk.  We
apply this scenario to the observed double BS-binary WOCS~7782 in the
old open cluster NGC~188, and show that its observed properties
\textbf{can be} reproduced within the context of the proposed model.
Based on this model, we predict the following properties for twin BSs:
(1) For the outer tertiary orbit, the initial orbital period should
lie between 220 days $\lesssim$ P$_{\rm out}$ $\lesssim$ 1100 days,
assuming initial masses for the inner binary components of $m_{\rm 1}
= 1.1$ M$_{\odot}$ and $m_{\rm 2} = 0.7$ to $0.9$M$_{\odot}$ and an
outer tertiary mass of $m_{\rm 3} = 1.4$ M$_{\odot}$.  \textbf{(2)}
After Roche-lobe overflow, the outer star turns into a white dwarf
(WD) of mass 0.43 to 0.54\,\MSun. There is a correlation between the
mass of this WD and the outer orbital period: more massive WDs will be
on wider orbits.  (3) The rotational axes of both BSs will be aligned
with each other and the orbital plane of the outer tertiary WD. (4)
The BSs will have roughly equal masses, independent of their initial
masses (since the lower mass star accretes the most).  The dominant
accretor should, therefore, be enriched more effectively by the
accreted material.  As a result, one of the BSs will appear to be more
enriched by either He, C and O or by s-process elements, depending on
whether the donor started to overflow its Roche lobe on, respectively,
the red giant or {\bf early} asymptotic giant branch.  (5) Relative to
old dense clusters with high-velocity dispersions, twin BSs in close
binaries formed from the proposed mechanism should be more frequent in
the Galactic field and younger open clusters with ages $\lesssim$ 4-6
Gyr, since then the donor will have a radiative envelope. (6) the
orbit of the binary BS will have a small semi-major axis (typically
$\aplt 0.3$\,au) and be close to circular ($e \aplt 0.1$).

@@

\textbf{Most} blue straggler stars are brighter and bluer than the main-sequence
(MS) turn-off in a cluster colour-magnitude diagram
\citep[e.g.][]{1953AJ.....58...61S,1989AJ.....98..217L,2014ApJ...782...49S}.
Two primary channels for BS formation have been proposed: mass
transfer from an evolved donor on to a MS star in a binary star system
\citep[e.g.][]{1964MNRAS.128..147M,1997A&A...328..143P,2009Natur.457..288K,2011MNRAS.410.2370L,2011Natur.478..356G},
and direct stellar collisions involving MS stars likely mediated via
binaries
\citep[e.g.][]{1975AJ.....80..809H,1997A&A...328..130P,2007ApJ...661..210L,2013MNRAS.428..897L,2013MNRAS.429.1221H, 2019A&A...621L..10P}.
\textbf{The first mechanism predicts BSs in binaries with WD companions,
whereas the second predicts MS companions in a wide and eccentric
binary.}  Other possible, albeit related, formation mechanisms include
mergers of close MS-MS binaries \citep{2019A&A...621L..10P}, and
mergers of the inner binaries of hierarchical triple star systems
induced by Lidov-Kozai oscillations coupled with tidal damping
\citep[e.g.][]{2009ApJ...697.1048P}.  \textbf{The latter predicts no binary companion, whereas the former predicts a MS companion in a wide binary.}

@@

Even more curious, there exists in the old OC NGC 188 a double BS
binary, called WOCS 7782 \citep{2009AJ....137.3743G}.  The BS
population in NGC 188 has a bi-modal period-eccentricity distribution.
As discussed in \citet{2011MNRAS.410.2370L}, this could be hinting at
a triple origin for at least some subset of the total BS population.
As for WOCS 7782, \citet{2009Natur.462.1032M} observed a compact and
mildly eccentric (i.e., $e \sim 0.1$) binary star system with an
orbital period of $\sim$ 10 days hosting two roughly equal-mass BSs.
During a given binary-binary interaction, the probability that not one
but two direct (MS-MS) collisions will occur is less than $10^{-2}$
\citep{1989AJ.....98..217L,2011MNRAS.410.2370L,2012MNRAS.425.2369L}.
Plus, binaries with collision products typically have relatively long
orbital periods \citep{2011Sci...334.1380F}. Dynamically, it is
difficult to form a short-period binary composed of two collision
products during a collisional interaction in a star cluster
\citep{2011MNRAS.410.2370L,2011Sci...334.1380F}, \textbf{and the
  timescale for exchanging another BS into a pre-existing BS-MS or
  BS-WD binary is much longer than the expected BS lifetime (see
  \citet{2011MNRAS.410.2370L} and the end of Section~\ref{sect:dyn}
  below)}.  So, how did WOCS 7782 form?

@@

In our scenario, we start with a binary star with component masses
$m_{\rm 1}$ and $m_{\rm 2}$ that is orbited by a tertiary of mass
$m_3$. The inner and outer binary orbital semi-major axes are denoted
a$_{\rm in}$ and a$_{\rm out}$, respectively.  \textbf{For
  specificty,} we assume both orbits, the inner as well as the outer,
\textbf{to be circular and in the same plane, which minimizes chaotic
  effects during the mass transfer process, facilitates more stable
  mass transfer and ultimately allows us to simulate our target system
  for longer while also maximizing the amount of mass transferred}.
These assumptions are also supported by the population of observed
low-mass triples \citep{2010yCat..73890925T,2018ApJ...854...44M}.
This initial configuration for our assumed formation scenario for WOCS
7782, described below, is depicted in figure~\ref{fig:fig1}.

@@

According to our scenario $m_3 > m_1 > m_2$ and the outer orbit is
sufficiently small that the tertiary star is filling its Roche lobe
and transfers mass to the inner binary before it {\bf reaches} the asymptotic
giant branch. We constrain the inner orbit by requiring the triple
system to be dynamically stable, for which we adopt eq.\,1 in
\cite{1999ASIC..522..385M}.  While transferring mass, the accretion
stream gathers around the inner binary at the circularization radius
a$_{\rm c}$, and forms a circumbinary disk
\citep{2002apa..book.....F}.  Using conservation of angular momentum,
we equate the specific angular momentum of the accreted mass at the
inner Lagrangian point of the (outer) donor star to the final specific
angular momentum of the accretion stream at the circularization radius
about the inner binary, this results in
\begin{equation}
\label{eqn:specangmom1}
v_{\rm orb,3}(a_{\rm out} - R_{\rm L}) = v_{\rm orb,c}a_{\rm c},
\end{equation}
where R$_{\rm L}$ is the radius of the Roche lobe of the outer
tertiary companion, $a_{\rm c}$ is the semi-major axis of the orbit
about the inner binary corresponding to the circularization radius and
v$_{\rm orb,c}$ is the orbital velocity at $a_{\rm c}$.  The distance
from the centre of mass corresponding to the tertiary defined by the
Roche lobe is given by eq.\,2 in \cite{1983ApJ...268..368E}.
Combining eq.\,2 in \citet{1983ApJ...268..368E} (with mass ratio q
$=$ m$_{\rm 3}$/(m$_{\rm 1} +$m$_{\rm 2}$)) with
eq.~\ref{eqn:specangmom1}, we solve for the circularization
radius as a function of a$_{\rm out}$ and the assumed stellar masses:
\begin{equation}
\label{eqn:ac}
a_{\rm c} = a_{\rm out}(1 - R_{\rm L}).
\end{equation}
In order for a circumbinary disk to form around the inner binary, we
require that $a_{\rm in} < a_{\rm c}$.

Figure~\ref{fig:fig2} shows the parameter space in the P$_{\rm
  out}$-P$_{\rm in}$-plane for WOCS 7782.  \textbf{Here we adopted,
  for clarity,} initial component masses of $m_{\rm 1} = 1.1$ M$_{\rm
  \odot}$ and $m_{\rm 2} = 0.9$ M$_{\rm \odot}$ for the inner binary
components, and $m_{\rm 3} = 1.4 $M$_{\rm \odot}$ for the outer
tertiary.  \textbf{The scenario worked out is general, but we opt for
  these specific parameters because they appear to naturally result in
  a system with parameters similar to WOCS 7782.}  We compare the
circularization radius to the semi-major axis of the inner binary, for
which we require $a_{\rm c} > a_{\rm in}$, after folding in all
constraints from the requirements for dynamical stability (listed in
the caption of figure~\ref{fig:fig2}), and the assumption of an outer
tertiary that is Roche lobe-filling \textbf{(see
  \citet{2014MNRAS.438.1909D} for more details)}.
%\simon{Why do you consider it important to mention that the orbit is hard, and for what environment did you consider this? - NL: Deleted.}  
Note that the range of plotted orbital periods P$_{\rm in}$
corresponding to a contact state for the inner binary lies outside the
range of plotted values for $P_{\rm in}$ (for components with radii of
1 R$_{\odot}$), since it does not contribute to constraining the outer
orbital properties.  The thick horizontal solid red line shows the
allowed range of outer semi-major axes, after folding in all of the
aforementioned criteria.  These constraints result in a rather narrow
range of initial conditions for the outer orbit, namely 2.2 $\times$
10$^{2}$ days $\le$ P$_{\rm out}$ $\le$ 1.1 $\times$ 10$^3$ days,
\textbf{which also directly translates into constraints on the final
  outer tertiary orbit.}
%, for
%which the tertiary star overfills its Roche lobe, the triple is (at
%least initially) dynamically stable and the accretion stream forms a
%gaseous ring around the inner binary.

\textbf{Finally, we compute the timescales for our hypothesized triple
  to undergo a direct interaction with another single or binary star.
  Using the same assumptions for the host cluster properties of NGC
  188 outlined in Section 3.2 of \citet{2011MNRAS.410.2370L}
  (right-hand column), we find upon setting the single-triple (3+1)
  and binary-triple (3+2) timescales equal to the expected duration of
  the mass transfer phase (i.e., $\sim 1$\, Myr) critical outer
  orbital periods for triples for $\gg 1$\,Myr.  These critical outer
  tertiary orbital periods, which correspond to the times for a
  specific triple to undergo an interaction, correspond to 3+1 and 3+2
  interaction times that are much longer than the maximum predicted
  outer orbital period of the hypothesized white dwarf tertiary in our
  scenario.  We therefore do not expect the mass transfer process to
  be interrupted by a dynamical interaction in the cluster center.}

@@

\begin{figure}[ht!]
  \includegraphics[width=\columnwidth]{fig_M14MSun_outerorbit.pdf}
  \caption{Giant radius as a function of the mass of the core of the
    Roche-lobe filling outer star (dark blue curve).  {\bf The first
      and last parts of this curve are orange to indicate that the
      star at these masses and radii is on the Hertzsprung-gap (to the
      left) or after core helium burning stage (to the right).  Only
      when the donor is on the giant branch (dark blue) Roche-lobe
      overflow will lead to a binary blue straggler.}  Here we adopt a
    donor mass of 1.4\,\MSun, but for an 1.2\,\MSun\, the donor the
    curve is quite similar (see dotted dark-blue curve).  The red
    squares in the curve show the parameters for which we performed
    more detailed gravitational-hydrodynamical simulations (see
    \S\,\ref{sims}).  The horizontal dashed line shows the orbital
    separation of the observed twin BS 7782.  The initial triple in
    which it possibly formed must at least have been dynamically
    stable. The minimal orbital separation for the inner binary for
    which the triple is stable is given by the lower green coloured
    curve.  Donors which are smaller than about 100\,\RSun\,
    (light-green curve indicated with a$_{\rm stable}$) result in a
    dynamically unstable triple. The minimal core mass associated with
    a stable triple is then indicated by the left-most vertical dotted
    line.  The orbital separation at which the donor star overfills
    its Roche lobe is indicated with the light-blue curve. The top
    curve (brown) shows an estimate of the final orbital separation of
    the outer star, and therefore of the final orbit of the WD around
    the inner twin BSs.  For core masses $\apgt 0.5$\,\MSun\, the
    final orbital separation, after mass transfer, is smaller than the
    initial orbit.  Here we adopted an initial inner binary mass of
    (1.0+0.9)\,\MSun\, and a final twin BS mass of (1.4+1.4)\,\MSun.
\label{fig:tertiarymass_vs_size}}
\end{figure}

@@

With the adopted parameters, we can also estimate the final orbital
period of the left-over core from the tertiary star after mass
transfer.  The change in orbital separation due to non-conservative
mass transfer can be expressed in terms of the mass of the outer star
before and after mass transfer, i.e. $m_3$ and $m'_3$ respectively,
the total mass in the inner binary before ($m_{\rm in}$) and after
accretion ($m'_{\rm in}$) and the amount of angular momentum lost per
unit mass $\eta \simeq 3$.  {\bf The value of $\eta = 3$ was derived
  in \cite{1991A&A...241..419P,1995A&A...296..691P} by matching the
  orbital evolution and birthrate of Be-type x-ray binaries that
  experience non-considervative mass transfer. This value of
  consistent with the analysis for non-conservative evolution of type
  B mass transfer in \cite{2011A&A...527A..84K} and further
  constrained in \cite{2007ASPC..367..387P} to understand the mass
  transfer in the 100-day orbital period binary V379 Cep.}  Adopting
the relation between the orbital separation before mass transfer ($a$)
and after mass transfer ($a'$) from \cite{1995A&A...296..691P}
\begin{equation}
  {a' \over a} = \left( {m_3 m_{\rm in} \over m'_3 m'_{\rm in}} \right)^{-2}
  \left( {m_3 + m_{\rm in} \over m'_3 + m'_{\rm in}} \right)^{2\eta + 1},
\end{equation}
we arrive at the top red curve in
figure\,\ref{fig:tertiarymass_vs_size}. This curve provides a prediction for 
the current orbital separation of the WD around the twin
BS 7782.
%% Tidal effects during mass \textbf{transfer}
%% have probably circularized the orbit, although some slight eccentricity
%% due to turbulent motion in the outer layers of the donor star may have
%% induced a small $e \ll 0.1$ eccentricity.

@@

We adopt initial masses of $m_{\rm 1} = 1.1$ M$_{\rm \odot}$ and
$m_{\rm 2} = 0.7$ M$_{\rm \odot}$ or 0.9\,M$_{\rm \odot}$ for the
inner binary components, and between $m_{\rm 3} = 1.2$ and $m_{\rm 3}
= 1.4$ M$_{\rm \odot}$ for the tertiary star.  We evolve the tertiary
star using the MESA stellar-evolution code \cite{2011ApJS..192....3P}
to a radius of about 100\,\RSun\, and 150\,\RSun, at which point we
assume it to overfill it's Roche lobe (see red \textbf{squares} in
figure~\ref{fig:tertiarymass_vs_size}).  We perform calculations for
an inner orbital separation of $a_{\rm in} = 0.10$\,au, $a_{\rm in} =
0.15$\,au and $a_{\rm in} = 0.20$\,au.  In total we performed 12
calculations at a resolution of 40k SPH particles and 12 at 80k.

@@

To test the \textbf{hypothesis that the} secondary in the inner binary
accretes more effectively than the primary star and to measure the
change to the inner orbit due to the Roche-lobe overflow of the outer
star, we perform a series of calculations in which we take the self
gravity and the hydrodynamical effects of the triple into account.
The results of \textbf{one of} these simulations (1091 days after
the onset of mass transfer) is presented in
figure~\ref{fig:mass_vs_semimajor_axis}.
%% figure~\ref{fig:topview_at_t1000day} and
%% The first figure
%% (figure~\ref{fig:topview_at_t1000day}) shows the top view of the same
%% initial realization for which we presented the initial conditions in
%% figure~\ref{fig:topview_at_t0} but now at an age of 1091 days after
%% the onset of mass transfer. We add, to the left panel,
%% the equipotential surfaces in the orbital plane.

It is apparent that the mass transfer in the adopted triples leads to
a rather untidy evolution, since much of the donor mass is lost
through the second Lagrangian point to the right side of the donor
star in figure~\ref{fig:topview_at_t1000day}. A considerable amount of
mass is also lost through the third Lagrangian point (to the left of
the inner binary), although it is hard to actually \textbf{quantify
  the amount of material lost}, becuase an appreciable fraction may to
rain back onto the triple system.  One remaining question is how much
mass is eventually ejected altogether from the triple system and is
therefore not accreted to any of the two inner stars. This value is
hard to estimate from the simulations, but an accretion efficiency of
$\apgt 0.8$ is necessary to make the scenario feasible. Over the time
scale for which we performed the calculations, this efficiency is
reached, but it is not clear how the system responds at later stages.

\begin{figure*}[ht!]
  \includegraphics[width=0.5\linewidth]{fig_BBSS_Lgas_M14Msun_A010au_t3360.png}
~  \includegraphics[width=0.5\linewidth]{fig_BBSS_gas_M14Msun_A010au_t3360.png}
  \caption{ Top view of one of the simulated triple systems at an age
    of $t \simeq 1091\,days$ after the start of the simulation when
    the 100\,\RSun\, outer star of 1.4\,\MSun\, over-fills its
    Roche-lobe.  The star is represented by 80000\,SPH particles and a
    core particle of $\sim 0.4$\,\MSun (black bullet in the middle of
    the right-most yellow blob). The inner binary \textbf{(to the
      left)} is represented by the yellow and red bullets for,
    respectively, the 1.1\,\MSun\, primary and 0.9\,\MSun\, secondary
    stars in a circular orbit of 0.1\,au. The 1.4\,\MSun\, giant star
    is presented to the right in a circular orbit with semi-major axis
    $\sim 250$\,\RSun\, in the plane of the inner binary.  Left panel
    shows the equipotential surfaces of the triple overplotted with
    the gas distribution, the right panel shows just the gas and the
    stars as bullets.
\label{fig:topview_at_t1000day}}
\end{figure*}

@@

In this paper, we propose a formation scenario for twin equal-mass
blue stragglers in tight binaries, as observed for WOCS 7782 in the
old OC NGC 188.  The proposed scenario involves mass transfer from an
evolved outer tertiary companion\textbf{. Part} of this mass is accreted by the
inner binary via a circumbinary disk\textbf{, while} the rest escapes through the
second and third Lagrangian points in the potential of the triple
system.  Our scenario makes several predictions for the observed
properties of a hypothetical outer triple companion, now a WD.  These
are:

>>>>

We propose a formation mechanism for twin blue stragglers (BSs) in
compact binaries that involves mass transfer from an evolved outer
tertiary companion on to the inner binary via a circumbinary disk.  We
apply this scenario to the observed double BS-binary WOCS~7782 in the
old open cluster NGC~188, and show that its observed properties
can be reproduced within the context of the proposed model.
Based on this model, we predict the following properties for twin BSs:
(1) For the outer tertiary orbit, the initial orbital period should
lie between 220 days $\lesssim$ P$_{\rm out}$ $\lesssim$ 1100 days,
assuming initial masses for the inner binary components of $m_{\rm 1}
= 1.1$ M$_{\odot}$ and $m_{\rm 2} = 0.7$ to $0.9$M$_{\odot}$ and an
outer tertiary mass of $m_{\rm 3} = 1.4$ M$_{\odot}$.  (2)
After Roche-lobe overflow, the outer star turns into a white dwarf
(WD) of mass 0.43 to 0.54\,\MSun. There is a correlation between the
mass of this WD and the outer orbital period: more massive WDs will be
on wider orbits.  (3) The rotational axes of both BSs will be aligned
with each other and the orbital plane of the outer tertiary WD. (4)
The BSs will have roughly equal masses, independent of their initial
masses (since the lower mass star accretes the most).  The dominant
accretor should, therefore, be enriched more effectively by the
accreted material.  As a result, one of the BSs will appear to be more
enriched by either He, C and O or by s-process elements, depending on
whether the donor started to overflow its Roche lobe on, respectively,
the red giant or early asymptotic giant branch.  (5) Relative to
old dense clusters with high-velocity dispersions, twin BSs in close
binaries formed from the proposed mechanism should be more frequent in
the Galactic field and younger open clusters with ages $\lesssim$ 4-6
Gyr, since then the donor will have a radiative envelope. (6) the
orbit of the binary BS will have a small semi-major axis (typically
$\aplt 0.3$\,au) and be close to circular ($e \aplt 0.1$).

@@

Most blue straggler stars are brighter and bluer than the main-sequence
(MS) turn-off in a cluster colour-magnitude diagram
\citep[e.g.][]{1953AJ.....58...61S,1989AJ.....98..217L,2014ApJ...782...49S}.
Two primary channels for BS formation have been proposed: mass
transfer from an evolved donor on to a MS star in a binary star system
\citep[e.g.][]{1964MNRAS.128..147M,1997A&A...328..143P,2009Natur.457..288K,2011MNRAS.410.2370L,2011Natur.478..356G},
and direct stellar collisions involving MS stars likely mediated via
binaries
\citep[e.g.][]{1975AJ.....80..809H,1997A&A...328..130P,2007ApJ...661..210L,2013MNRAS.428..897L,2013MNRAS.429.1221H, 2019A&A...621L..10P}.
The first mechanism predicts BSs in binaries with WD companions,
whereas the second predicts MS companions in a wide and eccentric
binary.  Other possible, albeit related, formation mechanisms include
mergers of close MS-MS binaries \citep{2019A&A...621L..10P}, and
mergers of the inner binaries of hierarchical triple star systems
induced by Lidov-Kozai oscillations coupled with tidal damping
\citep[e.g.][]{2009ApJ...697.1048P}.  The latter predicts no binary companion, whereas the former predicts a MS companion in a wide binary.

@@

Even more curious, there exists in the old OC NGC 188 a double BS
binary, called WOCS 7782 \citep{2009AJ....137.3743G}.  The BS
population in NGC 188 has a bi-modal period-eccentricity distribution.
As discussed in \citet{2011MNRAS.410.2370L}, this could be hinting at
a triple origin for at least some subset of the total BS population.
As for WOCS 7782, \citet{2009Natur.462.1032M} observed a compact and
mildly eccentric (i.e., $e \sim 0.1$) binary star system with an
orbital period of $\sim$ 10 days hosting two roughly equal-mass BSs.
During a given binary-binary interaction, the probability that not one
but two direct (MS-MS) collisions will occur is less than $10^{-2}$
\citep{1989AJ.....98..217L,2011MNRAS.410.2370L,2012MNRAS.425.2369L}.
Plus, binaries with collision products typically have relatively long
orbital periods \citep{2011Sci...334.1380F}. Dynamically, it is
difficult to form a short-period binary composed of two collision
products during a collisional interaction in a star cluster
\citep{2011MNRAS.410.2370L,2011Sci...334.1380F}, and the timescale for
exchanging another BS into a pre-existing BS-MS or BS-WD binary is
much longer than the expected BS lifetime (see
\citet{2011MNRAS.410.2370L} and the end of Section~\ref{sect:dyn}
below).  So, how did WOCS 7782 form?

@@

In our scenario, we start with a binary star with component masses
$m_{\rm 1}$ and $m_{\rm 2}$ that is orbited by a tertiary of mass
$m_3$. The inner and outer binary orbital semi-major axes are denoted
a$_{\rm in}$ and a$_{\rm out}$, respectively.  For specificty, we
assume both orbits, the inner as well as the outer, to be circular and
in the same plane, which minimizes chaotic effects during the mass
transfer process, facilitates more stable mass transfer and ultimately
allows us to simulate our target system for longer while also
maximizing the amount of mass transferred.  These assumptions are also
supported by the population of observed low-mass triples
\citep{2010yCat..73890925T,2018ApJ...854...44M}.  This initial
configuration for our assumed formation scenario for WOCS 7782,
described below, is depicted in figure~\ref{fig:fig1}.

@@

According to our scenario $m_3 > m_1 > m_2$ and the outer orbit is
sufficiently small that the tertiary star is filling its Roche lobe
and transfers mass to the inner binary before it reaches the asymptotic
giant branch. We constrain the inner orbit by requiring the triple
system to be dynamically stable, for which we adopt eq.\,1 in
\cite{1999ASIC..522..385M}.  While transferring mass, the accretion
stream gathers around the inner binary at the circularization radius
a$_{\rm c}$, and forms a circumbinary disk
\citep{2002apa..book.....F}.  Using conservation of angular momentum,
we equate the specific angular momentum of the accreted mass at the
inner Lagrangian point of the (outer) donor star to the final specific
angular momentum of the accretion stream at the circularization radius
about the inner binary, this results in
\begin{equation}
\label{eqn:specangmom1}
v_{\rm orb,3}(a_{\rm out} - R_{\rm L}) = v_{\rm orb,c}a_{\rm c},
\end{equation}
where R$_{\rm L}$ is the radius of the Roche lobe of the outer
tertiary companion, $a_{\rm c}$ is the semi-major axis of the orbit
about the inner binary corresponding to the circularization radius and
v$_{\rm orb,c}$ is the orbital velocity at $a_{\rm c}$.  The distance
from the centre of mass corresponding to the tertiary defined by the
Roche lobe is given by eq.\,2 in \cite{1983ApJ...268..368E}.
Combining eq.\,2 in \citet{1983ApJ...268..368E} (with mass ratio q
$=$ m$_{\rm 3}$/(m$_{\rm 1} +$m$_{\rm 2}$)) with
eq.~\ref{eqn:specangmom1}, we solve for the circularization
radius as a function of a$_{\rm out}$ and the assumed stellar masses:
\begin{equation}
\label{eqn:ac}
a_{\rm c} = a_{\rm out}(1 - R_{\rm L}).
\end{equation}
In order for a circumbinary disk to form around the inner binary, we
require that $a_{\rm in} < a_{\rm c}$.

Figure~\ref{fig:fig2} shows the parameter space in the P$_{\rm
  out}$-P$_{\rm in}$-plane for WOCS 7782.  Here we adopted, for
clarity, initial component masses of $m_{\rm 1} = 1.1$ M$_{\rm \odot}$
and $m_{\rm 2} = 0.9$ M$_{\rm \odot}$ for the inner binary components,
and $m_{\rm 3} = 1.4 $M$_{\rm \odot}$ for the outer tertiary.  The
scenario worked out is general, but we opt for these specific
parameters because they appear to naturally result in a system with
parameters similar to WOCS 7782.  We compare the circularization
radius to the semi-major axis of the inner binary, for which we
require $a_{\rm c} > a_{\rm in}$, after folding in all constraints
from the requirements for dynamical stability (listed in the caption
of figure~\ref{fig:fig2}), and the assumption of an outer tertiary
that is Roche lobe-filling (see \citet{2014MNRAS.438.1909D} for more
details).
%\simon{Why do you consider it important to mention that the orbit is hard, and for what environment did you consider this? - NL: Deleted.}  
Note that the range of plotted orbital periods P$_{\rm in}$
corresponding to a contact state for the inner binary lies outside the
range of plotted values for $P_{\rm in}$ (for components with radii of
1 R$_{\odot}$), since it does not contribute to constraining the outer
orbital properties.  The thick horizontal solid red line shows the
allowed range of outer semi-major axes, after folding in all of the
aforementioned criteria.  These constraints result in a rather narrow
range of initial conditions for the outer orbit, namely 2.2 $\times$
10$^{2}$ days $\le$ P$_{\rm out}$ $\le$ 1.1 $\times$ 10$^3$ days,
which also directly translates into constraints on the final
  outer tertiary orbit.
%, for
%which the tertiary star overfills its Roche lobe, the triple is (at
%least initially) dynamically stable and the accretion stream forms a
%gaseous ring around the inner binary.

Finally, we compute the timescales for our hypothesized triple to
undergo a direct interaction with another single or binary star.
Using the same assumptions for the host cluster properties of NGC 188
outlined in Section 3.2 of \citet{2011MNRAS.410.2370L} (right-hand
column), we find upon setting the single-triple (3+1) and
binary-triple (3+2) timescales equal to the expected duration of the
mass transfer phase (i.e., $\sim 1$\, Myr) critical outer orbital
periods for triples for $\gg 1$\,Myr.  These critical outer tertiary
orbital periods, which correspond to the times for a specific triple
to undergo an interaction, correspond to 3+1 and 3+2 interaction times
that are much longer than the maximum predicted outer orbital period
of the hypothesized white dwarf tertiary in our scenario.  We
therefore do not expect the mass transfer process to be interrupted by
a dynamical interaction in the cluster center.

@@

\begin{figure}[ht!]
  \includegraphics[width=\columnwidth]{fig_M14MSun_outerorbit.pdf}
  \caption{Giant radius as a function of the mass of the core of the
    Roche-lobe filling outer star (dark blue curve).  The first and
    last parts of this curve are orange to indicate that the star at
    these masses and radii is on the Hertzsprung-gap (to the left) or
    after core helium burning stage (to the right).  Only when the
    donor is on the giant branch (dark blue) Roche-lobe overflow will
    lead to a binary blue straggler.  Here we adopt a donor mass of
    1.4\,\MSun, but for an 1.2\,\MSun\, the donor the curve is quite
    similar (see dotted dark-blue curve).  The red squares in the
    curve show the parameters for which we performed more detailed
    gravitational-hydrodynamical simulations (see \S\,\ref{sims}).
    The horizontal dashed line shows the orbital separation of the
    observed twin BS 7782.  The initial triple in which it possibly
    formed must at least have been dynamically stable. The minimal
    orbital separation for the inner binary for which the triple is
    stable is given by the lower green coloured curve.  Donors which
    are smaller than about 100\,\RSun\, (light-green curve indicated
    with a$_{\rm stable}$) result in a dynamically unstable
    triple. The minimal core mass associated with a stable triple is
    then indicated by the left-most vertical dotted line.  The orbital
    separation at which the donor star overfills its Roche lobe is
    indicated with the light-blue curve. The top curve (brown) shows
    an estimate of the final orbital separation of the outer star, and
    therefore of the final orbit of the WD around the inner twin BSs.
    For core masses $\apgt 0.5$\,\MSun\, the final orbital separation,
    after mass transfer, is smaller than the initial orbit.  Here we
    adopted an initial inner binary mass of (1.0+0.9)\,\MSun\, and a
    final twin BS mass of (1.4+1.4)\,\MSun.
\label{fig:tertiarymass_vs_size}}
\end{figure}

@@

With the adopted parameters, we can also estimate the final orbital
period of the left-over core from the tertiary star after mass
transfer.  The change in orbital separation due to non-conservative
mass transfer can be expressed in terms of the mass of the outer star
before and after mass transfer, i.e. $m_3$ and $m'_3$ respectively,
the total mass in the inner binary before ($m_{\rm in}$) and after
accretion ($m'_{\rm in}$) and the amount of angular momentum lost per
unit mass $\eta \simeq 3$.  The value of $\eta = 3$ was derived
  in \cite{1991A&A...241..419P,1995A&A...296..691P} by matching the
  orbital evolution and birthrate of Be-type x-ray binaries that
  experience non-considervative mass transfer. This value of
  consistent with the analysis for non-conservative evolution of type
  B mass transfer in \cite{2011A&A...527A..84K} and further
  constrained in \cite{2007ASPC..367..387P} to understand the mass
  transfer in the 100-day orbital period binary V379 Cep.  Adopting
the relation between the orbital separation before mass transfer ($a$)
and after mass transfer ($a'$) from \cite{1995A&A...296..691P}
\begin{equation}
  {a' \over a} = \left( {m_3 m_{\rm in} \over m'_3 m'_{\rm in}} \right)^{-2}
  \left( {m_3 + m_{\rm in} \over m'_3 + m'_{\rm in}} \right)^{2\eta + 1},
\end{equation}
we arrive at the top red curve in
figure\,\ref{fig:tertiarymass_vs_size}. This curve provides a prediction for 
the current orbital separation of the WD around the twin
BS 7782.
%% Tidal effects during mass \textbf{transfer}
%% have probably circularized the orbit, although some slight eccentricity
%% due to turbulent motion in the outer layers of the donor star may have
%% induced a small $e \ll 0.1$ eccentricity.

@@

We adopt initial masses of $m_{\rm 1} = 1.1$ M$_{\rm \odot}$ and
$m_{\rm 2} = 0.7$ M$_{\rm \odot}$ or 0.9\,M$_{\rm \odot}$ for the
inner binary components, and between $m_{\rm 3} = 1.2$ and $m_{\rm 3}
= 1.4$ M$_{\rm \odot}$ for the tertiary star.  We evolve the tertiary
star using the MESA stellar-evolution code \cite{2011ApJS..192....3P}
to a radius of about 100\,\RSun\, and 150\,\RSun, at which point we
assume it to overfill it's Roche lobe (see red squares in
figure~\ref{fig:tertiarymass_vs_size}).  We perform calculations for
an inner orbital separation of $a_{\rm in} = 0.10$\,au, $a_{\rm in} =
0.15$\,au and $a_{\rm in} = 0.20$\,au.  In total we performed 12
calculations at a resolution of 40k SPH particles and 12 at 80k.

@@

To test the hypothesis that the secondary in the inner binary
accretes more effectively than the primary star and to measure the
change to the inner orbit due to the Roche-lobe overflow of the outer
star, we perform a series of calculations in which we take the self
gravity and the hydrodynamical effects of the triple into account.
The results of one of these simulations (1091 days after
the onset of mass transfer) is presented in
figure~\ref{fig:mass_vs_semimajor_axis}.
%% figure~\ref{fig:topview_at_t1000day} and
%% The first figure
%% (figure~\ref{fig:topview_at_t1000day}) shows the top view of the same
%% initial realization for which we presented the initial conditions in
%% figure~\ref{fig:topview_at_t0} but now at an age of 1091 days after
%% the onset of mass transfer. We add, to the left panel,
%% the equipotential surfaces in the orbital plane.

It is apparent that the mass transfer in the adopted triples leads to
a rather untidy evolution, since much of the donor mass is lost
through the second Lagrangian point to the right side of the donor
star in figure~\ref{fig:topview_at_t1000day}. A considerable amount of
mass is also lost through the third Lagrangian point (to the left of
the inner binary), although it is hard to actually quantify
  the amount of material lost, becuase an appreciable fraction may to
rain back onto the triple system.  One remaining question is how much
mass is eventually ejected altogether from the triple system and is
therefore not accreted to any of the two inner stars. This value is
hard to estimate from the simulations, but an accretion efficiency of
$\apgt 0.8$ is necessary to make the scenario feasible. Over the time
scale for which we performed the calculations, this efficiency is
reached, but it is not clear how the system responds at later stages.

\begin{figure*}[ht!]
  \includegraphics[width=0.5\linewidth]{fig_BBSS_Lgas_M14Msun_A010au_t3360.png}
~  \includegraphics[width=0.5\linewidth]{fig_BBSS_gas_M14Msun_A010au_t3360.png}
  \caption{ Top view of one of the simulated triple systems at an age
    of $t \simeq 1091\,days$ after the start of the simulation when
    the 100\,\RSun\, outer star of 1.4\,\MSun\, over-fills its
    Roche-lobe.  The star is represented by 80000\,SPH particles and a
    core particle of $\sim 0.4$\,\MSun (black bullet in the middle of
    the right-most yellow blob). The inner binary (to the
      left) is represented by the yellow and red bullets for,
    respectively, the 1.1\,\MSun\, primary and 0.9\,\MSun\, secondary
    stars in a circular orbit of 0.1\,au. The 1.4\,\MSun\, giant star
    is presented to the right in a circular orbit with semi-major axis
    $\sim 250$\,\RSun\, in the plane of the inner binary.  Left panel
    shows the equipotential surfaces of the triple overplotted with
    the gas distribution, the right panel shows just the gas and the
    stars as bullets.
\label{fig:topview_at_t1000day}}
\end{figure*}

@@

In this paper, we propose a formation scenario for twin equal-mass
blue stragglers in tight binaries, as observed for WOCS 7782 in the
old OC NGC 188.  The proposed scenario involves mass transfer from an
evolved outer tertiary companion. Part of this mass is accreted by the
inner binary via a circumbinary disk, while the rest escapes through the
second and third Lagrangian points in the potential of the triple
system.  Our scenario makes several predictions for the observed
properties of a hypothetical outer triple companion, now a WD.  These
are:
